### 2023-08-28 sections/1-introduction.tex
forgot to push a bunch of stuff so here we go

\section{Introduction}
\label{sec:introduction}
Often dubbed "the Big Data explosion," rapidly advancing technologies have enabled researchers to collect complex data on an unprecedented scale. 
This is especially true in the field of astronomy. 
The Sloan Digital Sky Survey (SDSS), for example, in its APOGEE dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions~\cite{alam2015eleventh}. 
Their MaNGA dataset also contains spectral measurements across the face of each of about 10,000 galaxies across 2,700 deg$^2$.

@@

In many fields, data collection rates outpace the rate of computing performance improvements predicted by Moore's Law~\cite{moore1965cramming}, indicating 
that computer architecture will not "catch up" to our computational needs in the near future. As a result, analyzing and learning from these large datasets 
requires better algorithms.

Many researchers would like the ability to perform similarity search on these datasets. 
Similatity search enables a variety of applications, including recommendation and classification systems. 
Additionally, with the rise of large language models such as GPT-3~\cite{brown2020language}, similarity search can become a useful tool for 
understanding and verifying the validity of model outputs. 
As the sizes and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become extremely challenging; 
even state-of-the-art algorithms exhibit a steep recall versus speed tradeoff~\cite{ishaq2019clustered}.

In particular, $k$-nearest neighbors (kNN) search is one of the most pervasive supervised classification methods in use~\cite{fix1952discriminatory, cover1967nearest}; most typically classification relies upon a majority vote among the $k$ nearest neighbors to a query.
Naive implementations of kNN, whose time complexity is linear in the dataset's cardinality, 
prove prohibitively slow for large datasets. While algorithms for fast kNN on large datasets exist, most of them are approximate and do not 
exploit the inherent structure and redundancy often present in big datasets. Approximate search is sufficient for many applications, but the 
need for efficient, \emph{exact} search remains.
In particular, while approximate kNN may agree with exact kNN in a majority vote for large values of $k$, it may be sensitive to local perturbations with smaller values of $k$.

Recent approaches to tackling the exponential growth of data include locality-sensitive hashing~\cite{bingham2001random}, 
clever indexing techniques such as the FM Index~\cite{ferragina2005indexing}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}. 
Entropy-scaling search is a paradigm for similarity search that exploits the inherent structure and redundancy often present in big datasets.
Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties (such as the 
\emph{metric entropy} and \emph{local fractal dimension}, defined in Section 2) of a dataset,
rather than its cardinality. In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest 
neighbors search to a hierarchical clustering approach. In this paper, we introduce CAKES (CLAM-Accelerated K-nearest-neighbor 
Entropy-scaling Search). CLAM (Clustered Learning of Approximate Manifolds), developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered} is a refinement of the CHESS clustering algorithm. Using the cluster tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform 
$k$-nearest neighbors search in addition to $\rho$-nearest neighbors search.

This paper focuses on improvements to CHESS, accelerated $\rho$- and $k$-nearest neighbors search, proof of a geometric guarantee for clusters, 
and the implementation of a variety of distance functions, including Euclidean, Manhattan, Cosine, Jaccard, Hamming, Levenshtein, Needleman-Wunch, 
and any user-defined function. We also provide a comparison of CAKES to FAISS~\cite{johnson2019billion}, FALCONN~\cite{andoni2015practical}, and
other algorithms in the ann-benchmarks suite~\cite{wang2017ann}.

>>>>

\section{Introduction}
\label{sec:introduction}
Often dubbed "the Big Data explosion," rapidly advancing technologies have enabled researchers to collect complex data at an unprecedented scale. 
This is especially true in the field of astronomy. 
The Sloan Digital Sky Survey (SDSS), for example, in its APOGEE dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions~\cite{alam2015eleventh}. 
Their MaNGA dataset also contains spectral measurements across the face of each of about 10,000 galaxies across 2,700 deg$^2$. [add sentence about big ANN competition]

@@

In many fields, the rate of data collection outpaces the rate of computing performance improvements predicted by Moore's Law~\cite{moore1965cramming}, indicating 
that computer architecture improvements will not "catch up" to our computational needs in the near future. As a result, analysis of these large datasets 
requires better algorithms.

Many researchers are especially interested in performing similarity search on these datasets. 
Similarity search enables a variety of applications, including recommendation and classification systems. 
Additionally, with the rise of large language models such as GPT-3~\cite{brown2020language}, similarity search can become a useful tool for 
understanding and verifying the validity of model outputs. 
As the sizes and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become extremely challenging; 
even state-of-the-art algorithms exhibit a steep recall versus speed tradeoff~\cite{ishaq2019clustered}.

In particular, $k$-nearest neighbors ($k$-NN) search is one of the most pervasive classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}; 
Naive implementations of kNN, whose time complexity is linear in the dataset's cardinality, 
prove prohibitively slow for large datasets. While algorithms for fast $k$-NN on large datasets exist, most of them are approximate and do not 
exploit the geometric and topological structure inherent in big datasets. Approximate search is sufficient for some applications, but the 
need for efficient, \emph{exact} search remains.
In particular, while approximate $k$-NN may agree with exact $k$-NN in a majority vote for large values of $k$, it may be sensitive to local perturbations with smaller values of $k$.

Recent approaches to tackling the exponential growth of data include locality-sensitive hashing~\cite{bingham2001random}, 
clever indexing techniques such as the FM Index~\cite{ferragina2005indexing}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}. 
Entropy-scaling search is a paradigm for similarity search that exploits the geometric and topological structure inherent in big datasets.
Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties (such as the 
\emph{metric entropy} and \emph{local fractal dimension}, defined in Section 2) of a dataset,
rather than its cardinality. In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest 
neighbors search to a hierarchical clustering approach. In this paper, we introduce CAKES (CLAM-Accelerated $K$-nearest-neighbor 
Entropy-scaling Search). CLAM (Clustered Learning of Approximate Manifolds), developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the CHESS clustering algorithm. 
Using the cluster tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform 
$k$-nearest neighbors search in addition to $\rho$-nearest neighbors search (also known as ranged nearest neighbors search).

This paper focuses on accelerated $\rho$- and $k$-nearest neighbors search. We explain some 
implementation improvements to CHESS, as well as improved genericity across distance functions. 
We also include proof of some geometric guarantees for clusters. Finally, we provide a comparison of 
CAKES to the state-of-the-art in similarity search, FAISS~\cite{johnson2019billion} on several datasets 
from the ann and big ann benchmarks suites.

### 2023-08-28 sections/2-methods.tex
forgot to push a bunch of stuff so here we go

In this manuscript, we are primarily concerned with $k$-nearest neighbors search in a finite-dimensional 
vector space. Some of the algorithms we present use a modified version of the $\rho$-nearest neighbors search 
algorithm described in CHESS~\cite{ishaq2019clustered} to conduct $k$-NN search. (TODO: make second sentence less 
vague once we determine exactly what we're doing with sharding and indexing.)

@@

We define a \emph{distance function} $f : (\textbf{X}, \textbf{X}) \mapsto \mathbb{R}^+ \cup \{0\}$ as a function which, 
given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number. We require that the distance function 
be symmetric ($f(x, y) = f(y, x)$ for all $x, y \in \textbf{X}$) and that the distance between two points $x$ and $y$ be zero if and only if $x = y$. 
When, in addition to these constraints, the distance function obeys
the triangle inequality, it is a \emph{distance metric}, and in such cases we can guarantee that our search algorithms have perfect recall. 
Choice of distance function varies by type of data. For example, with electromagnetic spectra, we use both 
Euclidean (L2) and Cosine distances. With biological or string data on the other hand, Hamming and Levenshtein distances are more appropriate.

Some of our algorithms for $\rho$- and $k$-nearest neighbors search rely on the \emph{local fractal dimension} at some point in the dataset, 
which we define as: 
$$\text{log}_2 \left( \frac{|B_X(q, r_1)|}{|B_X(q, r_2)|} \right)$$
where $B_X(q, r)$ is the set of points contained in a ball of radius $r$ 
centered at a point $q$ in the dataset $\textbf{X}$; in this example, we compute fractal dimension for some radius $r_1$ and a smaller radius $r_2 = \frac{r_1}{2}$.
We stress that this concept differs from the \emph{embedding dimension} of a dataset. To illustrate the difference,
consider SDSS's APOGEE dataset, wherein each datum is a nonnegative real-valued vector of length 8,575. Hence, the \emph{embedding dimension} of this dataset is 8,575. 
However, due to physical constraints (namely, the laws that govern stellar fusion and spectral emission lines), the data are constrained to a lower-dimensional 
manifold within the 8,575-dimensional embedding space. The \emph{local fractal dimension} is an approximation of the dimensionality of that lower-dimensional manifold at a given point, for some length scale.
The notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold is known as the \emph{manifold hypothesis}.

We define a \emph{cluster} as a set of points with a \emph{center}, a \emph{radius}, and an approximation of the \emph{local fractal dimension}.
The \emph{center} is the geometric median of the set of points in the \emph{cluster}, and so it is a real data point. The \emph{radius} is the
maximum distance from a point in the cluster to its \emph{center}. We estimate \emph{local fractal dimension} at the cluster radius and half
the cluster radius. Each cluster (unless it is a leaf cluster) has two child clusters in much the same way that a node in
a binary tree has two child nodes. We define the \emph{metric entropy} of a data set under a hierarchical clustering scheme as a refinement of [7], where
metric entropy for a given cluster radius $r_c$ was the number of clusters of radius $r_c$ needed to cover all data. Here, we use a binary, divisive clustering 
approach, but with early stopping criteria; clusters with fewer than a threshold number of points are not further split. 
Since we frame the asymptotic complexity of $\rho$-NN search in terms of the number of leaf clusters, the \emph{metric entropy} is best thought of in terms of the number of
leaf clusters.

With these concepts defined, we can now pose the $k$- and $\rho$ nearest neighbors search problems.
Given a query $q$, along with a distance function $f$ defined on a dataset $ \textbf{X}$, $k$-nearest neighbors search aims to: 
for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$; that is, find the $k$ points $x \in \textbf{X}$ with the smallest $f(q, x)$.
We also have the $\rho$-nearest neighbors search problem: for a given radius $\rho$, find all points in $\textbf{X}$ that are a distance of $\rho$ 
or less from $q$; that is, all points $x \in \textbf{X}$ such that $f(q, x) \leq \rho$.

@@

We sometimes also use this notation with points instead of clusters. For example, we sometimes use $\delta = f(x, q)$ to refer to the distance 
from a point $x$ to the query $q$. Since we can think of a point as having radius $r=0$, for points we have that $\delta_{max} = \delta = \delta_{min}$.

@@

We start by building a divisive hierarchical clustering of the dataset with CLAM, using a 
similar recursive procedure as outlined in CHESS~\cite{ishaq2019clustered}, but with the following 
improvements: better selection of poles for partitioning and memoization of useful cluster properties, 
such as local fractal dimension (see Section 2.4.2: $k$-NN by Repeated $\rho$-NN).

@@

For a cluster with $m$ points, we begin by computing a 
random sample of $\sqrt m$ points. We then compute the geometric median of this sample, which we call the 
cluster's center. We then compute the distance from the center to each point in the sample of $\sqrt m$ points. 
The point $l$ which is furthest from the center is designated the left pole, and the point $r$ which is furthest
from $l$ is designated the right pole. We then partition the cluster into a left child and right child, where the 
left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all 
points in the cluster which are closer to $r$ than to $l$. Without loss of generality, we assign to the left child 
those points which are equidistant from $l$ and $r$. Starting from a root-cluster containing the entire dataset, we 
repeat this procedure until each leaf contains only one datum or until some other user-specified stopping criterion 
is met.

@@

@@

\subsection {Index-Building}
\begin{itemize}
    \item sharded search
    \item dataset depth-first reordering 
    \item index building 
    \item optimization of number of shards 
\end{itemize}

@@

We conduct $\rho$-nearest neighbors search similarly as described in CHESS~\cite{ishaq2019clustered}, but 
with some implementation improvements including identification of candidate clusters which are leaves and 
conducting exhaustive search over those leaves.

@@

In this section, we present four novel variants of $k$-nearest neighbors search: Greedy Search, $k$-NN by Repeated $\rho$-NN, Sieve Search I, and Sieve Search II. 
We use a preliminary auto-tuning function (TODO: do we though?) to predict the variant which will perform 
best for a given query, dataset, and value of $k$. We then proceed with search using that variant.

\subsubsection{Greedy Search}
\label{subsubsec:methods:knn-search:greedy-search}
Greedy Search repeatedly partitions the cluster with the smallest $\delta_{min}$ into its children 
until the cluster with the smallest $\delta_{min}$ is a leaf. We then remove it and add its points to a fixed size priority queue of points. 
Search terminates when the furthest hit in the queue of points has $\delta$ greater than the $\delta_{min}$ of the closest remaining cluster.

@@

While $P$ has fewer than $k$ clusters or $\delta$ of the top priority element in $P$ is greater 
than $\delta_{min}$ of the top priority element in $Q$, do the following:
\begin{enumerate}
\item While the top priority element $q$ in $Q$ is not a leaf, replace $q$ with its children.
\item Remove $q$ from $Q$ and add all of its points to $P$. 
\item Remove points from $P$ until $P$ has $k$ points. 
\end{enumerate}
After this process, the points left in $P$ are the $k$ nearest neighbors of $q$.

@@

\subsubsection{$k$-NN by Repeated $\rho$-NN}
\label{subsubsec:methods:knn-search:repeated-rnn}
In this $k$-NN variant, we repeatedly perform
$\rho$-nearest neighbors search, increasing the search radius until $k$ neighbors
have been found.

Let $P$ be the set of nearest neighbors found thus far.
Search starts with a radius $m$ equal to the radius of the cluster tree divided by
the cardinality of the dataset. We perform $\rho$-NN search with radius $m$. 
If no points are within a distance $m$ of the query, we increase the radius by a factor of 
2 and perform $\rho$-NN search again, repeating until at least one point is found, i.e., 
until $|P| > 0$.

Once $|P| > 0$, we continue to perform $\rho$-NN search, but instead of 
increasing the radius by a factor of 2 on each iteration, we increase it by a factor determined 
by the local fractal dimension of the clusters containing the points found so far. In particular, 
we increase the radius by a factor of 
$$\text{min}\left(2, \frac{k}{|P|^{\frac{1}{\mu}}}\right)$$
where $\mu$ is the mean local fractal dimension (abbreviated lfd in algorithm below) of the clusters containing the points found so far.
Intuitively, the factor by which we increase the radius should be \emph{inversely} proportional to the number of points found so far. 
Additionally, when the local fractal dimension near the query and found neighbors is low, this suggests that the data 
are relatively concentrated in that region; thus, the factor of radius increase should be \emph{directly} proportional to the 
local fractal dimension. Since the local fractal dimension is memoized at clustering time, this poses no additional cost at search time.
Once $|P| >= k$, we sort the points in $P$ by $\delta$ and return the $k$ closest points.

@@

\subsubsection{Sieve Search I}
\label{subsubsec:methods:knn-search:sieve}
With Sieve Search I, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{min}$ greater than $\tau$ 
can contain one of the $k$ nearest neighbors. Search terminates when we have a threshold containing exactly $k$ points.

@@

\subsubsection{Sieve Search II}
\label{subsubsec:methods:knn-search:sieve2}

\subsubsection{Linear Search}
\label{subsubsec:methods:knn-search:leaf-search}

Let $hits$ be a a fixed size priority queue with $k$ instances sorted by non-decreasing $f(q, x)$, i.e. distance to the query.
Initialize $hits$ with the centers of $candidates$.

Let $f_k$ be the distance to the $k^{th}$ farthest instance so far.
Exclude all $candidates$ for which $\delta^2 > f_k$, i.e. their closest possible instance is farthest that the $k^{th}$ farthest instance found so far.
Remove the closest cluster from $candidates$ and add all its contained instances to $hits$.
Repeat until $candidates$ is empty.

\subsubsection{Complexity}
\label{subsubsec:methods:knn-search:complexity}

This is at-least on-par with the complexity for $\rho$-nearest neighbors search from CHESS.
It is potentially better because we effectively shrink the search radius to adapt to the $k$ closest instances found during the algorithm.

>>>>

In this manuscript, we are primarily concerned with $k$-nearest neighbors in a finite-dimensional 
vector space.

@@

We define a \emph{distance function} $f : (\textbf{X}, \textbf{X}) \mapsto \mathbb{R}^+ \cup \{0\}$ as a function which, 
given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number. We require that the distance function 
be symmetric ($f(x, y) = f(y, x)$ for all $x, y \in \textbf{X}$) and that the distance between two points $x$ and $y$ be zero if and only if $x = y$. 
When, in addition to these constraints, the distance function obeys
the triangle inequality, it is a \emph{distance metric}, and in such cases we can guarantee that our search algorithms have perfect recall. 
Choice of distance function varies by type of data. For example, with electromagnetic spectra, we use both 
Euclidean (L2) and Cosine distances. With biological or string data, Hamming and Levenshtein distances are more appropriate.

Some of our algorithms for $\rho$- and $k$-nearest neighbors search rely on the \emph{local fractal dimension} at some point in the dataset, 
which we define as: 
$$\text{log}_2 \left( \frac{|B_X(q, r_1)|}{|B_X(q, r_2)|} \right)$$
where $B_X(q, r)$ is the set of points contained in a ball of radius $r$ 
centered at a point $q$ in the dataset $\textbf{X}$; in this example, we compute fractal dimension for some radius $r_1$ and a smaller radius $r_2 = \frac{r_1}{2}$.
We stress that this concept differs from the \emph{embedding dimension} of a dataset. To illustrate the difference,
consider SDSS's APOGEE dataset, wherein each datum is a nonnegative real-valued vector of length 8,575. Hence, the \emph{embedding dimension} of this dataset is 8,575. 
However, due to physical constraints (namely, the laws that govern stellar fusion and spectral emission lines), the data are constrained to a lower-dimensional 
manifold within the 8,575-dimensional embedding space. The \emph{local fractal dimension} is an approximation of the dimensionality of that lower-dimensional manifold at a given point, for some length scale.
The notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold is known as the \emph{manifold hypothesis}.
[TODO: add example for GreenGenes and add lfd plots from CHESS]

We define a \emph{cluster} as a set of points with a \emph{center}, a \emph{radius}, and an approximation of the \emph{local fractal dimension}.
The \emph{center} is the geometric median of a sample of points in the \emph{cluster}, and so it is a real data point. The \emph{radius} is the
maximum distance from a point in the cluster to its \emph{center}. We estimate \emph{local fractal dimension} at the cluster radius and half
the cluster radius. Each cluster (unless it is a leaf cluster) has two child clusters in much the same way that a node in
a binary tree has two child nodes. We define the \emph{metric entropy} of a data set under a hierarchical clustering scheme as a refinement of [7], where
metric entropy for a given cluster radius $r_c$ was the number of clusters of radius $r_c$ needed to cover all data. Here, we use a hierarchical, divisive clustering 
approach, but with early stopping criteria; clusters which satisfy a user-specified stopping criterion (e.g. a specified cardinality or radius) are not further split. 
Since we frame the asymptotic complexity of $\rho$-NN search in terms of the number of leaf clusters, the \emph{metric entropy} is best thought of in terms of the number of
leaf clusters.

[TODO: Do set comprehension definitions for these and then prose refinement]
With these concepts defined, we can now pose the $k$- and $\rho$ nearest neighbors search problems.
Given a query $q$, along with a distance function $f$ defined on a dataset $ \textbf{X}$, $k$-nearest neighbors search aims to: 
for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$; that is, find the $k$ points $x \in \textbf{X}$ with the smallest $f(q, x)$.
We also have the $\rho$-nearest neighbors search problem: for a given radius $\rho$, find all points in $\textbf{X}$ that are a distance of $\rho$ 
or less from $q$; that is, all points $x \in \textbf{X}$ such that $f(q, x) \leq \rho$.

@@

[TODO: talk about singleltons or whatever and overloading of notation]
We sometimes also use this notation with points instead of clusters. For example, we sometimes use $\delta = f(x, q)$ to refer to the distance 
from a point $x$ to the query $q$. Since we can think of a point as having radius $r=0$, for points we have that $\delta_{max} = \delta = \delta_{min}$.

@@

[TODO: take out memoization and talk about reordering instead]
We start by building a divisive hierarchical clustering of the dataset with CLAM, using a 
similar recursive procedure as outlined in CHESS~\cite{ishaq2019clustered}, but with the following 
improvements: better selection of poles for partitioning and memoization of useful cluster properties, 
such as local fractal dimension (see Section 2.4.2: $k$-NN by Repeated $\rho$-NN).

@@

[TODO: use cluster C and |C| instead here; make sure you mention how radius is calculated; sampling, center, radius, poles; 
define geometric median]
For a cluster with $m$ points, we begin by computing a 
random sample of $\sqrt m$ points. We then compute the geometric median of this sample, which we call the 
cluster's center. We then compute the distance from the center to each point in the sample of $\sqrt m$ points. 
The point $l$ which is furthest from the center is designated the left pole, and the point $r$ which is furthest
from $l$ is designated the right pole. We then partition the cluster into a left child and right child, where the 
left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all 
points in the cluster which are closer to $r$ than to $l$. Without loss of generality, we assign to the left child 
those points which are equidistant from $l$ and $r$. Starting from a root-cluster containing the entire dataset, we 
repeat this procedure until each leaf contains only one datum or until some other user-specified stopping criterion 
is met.

@@

\subsubsection {Dataset Reordering}

[2 tiers of improvement]

CLAM-CAKES also reorders the dataset to improve the tree-building process 
used in CHESS. In CHESS, each cluster stored a vector of indices into the dataset, which 
were used to retrieve its points during search. Thus, for a dataset with cardinality $n$, 
at each depth in the tree, we would store a total of $n$ indices. Since we have approximately

@@

\subsection {Index-Building}
\begin{itemize}
    \item sharded search
    \item index building 
    \item optimization of number of shards 
\end{itemize}

@@

[TODO: delete this and talk about overlapping children improvement instead]
We conduct $\rho$-nearest neighbors search similarly as described in CHESS~\cite{ishaq2019clustered}, but 
with some implementation improvements including identification of candidate clusters which are leaves and 
conducting exhaustive search over those leaves.

@@

[TODO: reorder so rho nn is first, then sieves, then Greedy Sieve]
In this section, we present four novel algorithms for $k$-nearest neighbors search: Greedy Search, $k$-NN by Repeated $\rho$-NN, Sieve Search I, and Sieve Search II. 
We also have an implementation of linear search. We use a preliminary auto-tuning function to predict the variant which will perform 
best for a given query, dataset, and value of $k$. We then proceed with search using that variant.

[TODO: tighten up or omit those opening paragraphs which roughly explain the algorithm]

[TODO: call queues of points H]
\subsubsection{Greedy Search}
\label{subsubsec:methods:knn-search:greedy-search}
Greedy Search repeatedly partitions the cluster with the smallest $\delta_{min}$ into its children 
until the cluster with the smallest $\delta_{min}$ is a leaf. We then remove it and add its points to a fixed size priority queue of points. 
Search terminates when the furthest hit in the queue of points has $\delta$ greater than the $\delta_{min}$ of the closest remaining cluster.

@@

[TODO: replace with pop and push language and start with initial pop]
While $P$ has fewer than $k$ clusters or $\delta$ of the top priority element in $P$ is greater 
than $\delta_{min}$ of the top priority element in $Q$, do the following:
\begin{enumerate}
\item While the top priority element $q$ in $Q$ is not a leaf, replace $q$ with its children.
\item Remove $q$ from $Q$ and add all of its points to $P$. 
\item Remove points from $P$ until $P$ has $k$ points. 
\end{enumerate}
After this process, the points left in $P$ are the $k$ nearest neighbors of $q$.

[TODO: fix to use priority queue language]

@@

\subsubsection{$k$-NN by Repeated $\rho$-NN}
\label{subsubsec:methods:knn-search:repeated-rnn}
In this algorithm, we perform
$\rho$-nearest neighbors search at [TODO: add radius], repeatedly increasing the search radius until $k$ neighbors
have been found.

[TODO: change P to H]
Let $P$ be the set of nearest neighbors found thus far.
Search starts with a radius $m$ equal to the radius of the cluster tree divided by
the cardinality of the dataset. We perform $\rho$-NN search with radius $m$. 
If no points are within a distance $m$ of the query, we increase the radius by a factor of 
2 and perform $\rho$-NN search again, repeating until at least one point is found, i.e., 
until $|P| > 0$.

[TODO: Add number next to all equations and fix factor so that it is correct (k should be raised too)]
Once $|P| > 0$, we continue to perform $\rho$-NN search, but instead of 
increasing the radius by a factor of 2 on each iteration, we increase it by a factor determined 
by the local fractal dimension of the clusters containing the points found so far. In particular, 
we increase the radius by a factor of 
$$\text{min}\left(2, \frac{k}{|P|^{\frac{1}{\mu}}}\right)$$
where $\mu$ is the mean local fractal dimension (abbreviated lfd in algorithm below) of the clusters containing the points found so far.
Intuitively, the factor by which we increase the radius should be \emph{inversely} proportional to the number of points found so far. 
Additionally, when the local fractal dimension near the query and found neighbors is low, this suggests that the data 
are relatively concentrated in that region; thus, the factor of radius increase should be \emph{directly} proportional to the 
local fractal dimension. Since the local fractal dimension is memoized at clustering time, this poses no additional cost at search time.
Once $|P| >= k$, we sort the points in $P$ by $\delta$ and return the $k$ closest points.

@@

[TODO: change sieve II to sieev with separate centers]
\subsubsection{Sieve Search I}
\label{subsubsec:methods:knn-search:sieve}
[TODO: make writeup here tighter and more consistent with what's in the algorithm; talk 
about enum variants here]
With Sieve Search I, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{min}$ greater than $\tau$ 
can contain one of the $k$ nearest neighbors. Search terminates when we have a threshold containing exactly $k$ points.

@@

\subsubsection{Sieve Search II}
\label{subsubsec:methods:knn-search:sieve2}
Sieve Search II is exactly the same as Sieve Search I, but with the following modification: clusters 
in $Q$ are represented twice-- once as their center and once as the rest of their points. 
Because we have that for any cluster $C$, $C.\delta \leq C.\delta_{max}$, representing the cluster 
center separately from the rest of the points causes the threshold $\tau$ to decrease more quickly, 
thus allowing us to eliminate some clusters from $Q$ earlier in the process.

\subsubsection{Complexity of $k$-NN Search}
\label{paragraph:methods:knn-complexity}
Add asymptotic complexity of various $k$-NN algorithms here.

### 2023-08-28 sections/5-discussion.tex
forgot to push a bunch of stuff so here we go

Is it the case that the optimal clusters for compression are also the optimal clusters for accelerated search?

What kind of trade-offs can we offer for space-savings of compression vs time-savings of accelerated-search in the compressed space?

@@

@@

Better methods for constructing minimal encodings, perhaps using domain expertise.

>>>>

@@

compression

@@
